Emergence of $^4$H $J^\pi=1^-$ resonance in contact theories

### 2022-06-18 manuscript/Draft_2.1.tex
from hexa

\author{Johannes Kirscher}
\address{Department of Physics, SRM University - AP, Amaravati 522502, Andhra Pradesh, India}

@@

Elastic neutron-triton scattering is considered at leading order in pionless effective field theory.
In the infinite-cutoff limit, we obtain the scattering length $a_0(\infty) = 3.06(1)$~fm
and the effective range $r_0(\infty) = 1.59(7)$~fm ($S$-wave). 
In the $P$-wave channel, we predict a scattering volume $a_1(\infty) = -9.8(1.6)$~fm$^3$
and a range equivalent $r_1(\infty) = 1.3(4)$~fm$^{-1}$ with the same zero-range theory. 
%
The results are extracted with three numerical techniques: a
Gaussian-stochastic-variational method, an integration of the configuration-space Faddeev-Yakubovsky 
equations, and a resonating-group reduction of the four-body problem to an effective two-body cluster theory. 
Renormalization-group dependence is assessed through variation of a cutoff regulator parametrized in a
range between $1\,\text{fm}^{-1}$ and $10\,\text{fm}^{-1}$.
%
Most remarkably, we find a cutoff-stable/RG-invariant resonance in the $^4$H $J^\pi=1^-$ system.
This $P$-wave resonance is thus a universal consequence of the unitary deuteron system
and an ensuing discretely scale invariant three-body system with a scale set by the triton binding energy.
%
This stabilization of a resonant state in a few-fermion system solely through pure contact interactions
has a significant consequences for the powercounting of the pionless theory.
Specifically, the existence of a resonance implies that the known inability of the theory's leading order
to bind larger nuclei like 16-oxygen can be cured in perturbation theory.
\end{abstract}

@@

Contact effective field theory (EFT) is not only practical in describing
bosonic~\cite{Bazak:2016wxm,Carlson:2017txq,Bazak:2018qnu,Bedaque:1998kg,Bedaque:1998km} 
and few-nucleon~\cite{Bedaque:1999ve,Barnea:2013uqa} systems, it represents a systematic
expansion of the underlying scales and their interactions~\cite{Hammer:2019poc,Barnea:2013uqa}, and
thereby provides a transparent framework to study the emergence of many-body phenomena from few-body properties.
However, the nuclear incarnation of such a theory, Pionless effective field theory (\eftnopi), apparently fails to predict
stable bound states for $P$-shell
nuclei\footnote{Fermi systems with more particles than accessible internal states}
at the leading order (LO)~\cite{Stetcu:2006ey,Contessi:2017rww,Dawkins:2019vcr,Schafer:2020ivj}. 
%
For instance, the lightest of such nuclei, $^6 {\rm Li}$, is 
found unstable with respect to breakup into $\rm ^4He$ in its ground state ($\alpha$)
and a deuteron ($d$) in the zero-range contact limit~\cite{Schafer:2020ivj}~of LO~\eftnopi.
Whether or not the theory finds the physical bound state at LO veiled as a virtual or resonant state which
reveals its physical character at higher orders is arguably the most imminent problem of the development of contact EFTs for
Fermi systems.
As perturbative higher-order insertions cannot create poles {\it ex novo}, the conceivable absence of such poles
within the convergence radius of LO~\eftnopi~would make the theory in its current version useless for the description
of larger nuclei.
The challenge cast in simple words: Does the LO~\eftnopi~support any type of non-zero poles at complex momenta below its
breakdown scale?

Historically, resonances were studied in (A$\le$4)-body bosonic systems close to the unitary limit\cite{Deltuva:2012ig,deltuva2020energies} confirming the existence of a tower of Efimov resonances. 
A recent study\cite{Habashi:2020qgw} proved the possibility of creating two-body $S$-shell resonances in the contact limit with negative effective range.%, naidon2017efimov}. 
In the \eftnopi at LO framework, a resonance have been also found in neutral hypernuclear system $\Lambda nn$ \cite{Schafer:2020rba} and $S$-wave virtual states have been confirmed as exited states of triton \cite{Rupak:2019} and hypertriton \cite{Schafer:2020rba}. 
Finally, a recent study of $^3n$ confirmed the absence of $P$-wave resonance or virtual state in this system\cite{Dietz:2021haj}. 
At the present knowledge of the authors, the calculations of $P$-shell systems stops at this level and no studies have been done for discrete invariant systems of equal mass.
The major goal of this work is to study resonances in $^4$H a system never addressed before in this framework and that can be understood starting from its universal behavior. 
The existence of an eventual $P$-shell resonance using \eftnopi at LO would open the possibility of finding the same structure also in heavier nuclei, with the chance of stabilizing many-nucleon systems by perturbative corrections.
%

%
Experimentally, $^4$H resonances are clearly seen in n-$^3$H elastic cross sections \cite{PhysRev.119.1981,SEAGRAVE1972250}, as well as in $\pi^-$ absorption experiments \cite{Sennhauser:1981tr,Gornov:1991fg}, or in transfer reactions \cite{Miljanic:1986zz,Sidorchuk:2003fwa}. 
However, extracted resonance parameters differ substantially, depending on the used protocol and on the underlying hypothesis in the data analysis. 
For instance, results of the same $\pi^-$+$^9$Be$\,\to\,$d+t+$^4$H experiment, performed by the same authors, yields for the lowest state ($E_R$=3.0$\pm$0.2~MeV, $\Gamma$=4.7$\pm$1~MeV) \cite{Gornov:1991fg}, ($E_R$=2.0$\pm$0.2~MeV, $\Gamma$=1.2$\pm$0.2~MeV) \cite{Gurov:2005sp}, and ($E_R$=1.6 $\pm$0.1 MeV, $\gamma^2$=0.4$\pm$0.1~MeV)\cite{Gurov2005p}, where $\gamma$ is the reduced width (see the original paper), depending whether one assumes the experimental signal to be created respectively by one, two, or three resonant states.
Independently, the $R$-matrix analysis of n-$^3$H cross section \cite{Tilley:1992zz} concluded about a superposition of several $P$-waves resonant states with quantum numbers $J^\pi$=$0^-$,$1^-$,$2^-$. 
With two non-degenerate $J^\pi$=$1^-$ states (denoted $1^-_I$ and $1^-_{II}$ ) corresponding to two different total spin contributions in the asymptotic channels\cite{Lazauskas:2004uq}. 
The lowest state parameters found in this case ($E_R$=3.2 MeV, $\Gamma$=5.4 MeV) are hardly compatible with the previous many-state analysis. 
The discrepancies among different experimental approaches further
increase when including the transfer reaction experiments.
%

From a numerical point of view, computing the $S$-matrix pole of a resonance is not an easy task. 
Microscopic ab initio calculations for nuclear few-body systems are therefore not abundant and up to now have been performed for a limited amount of particles ($A<6$) \cite{Li:2019pmg,Lazauskas:2019cxj}. 
This limit is often circumvented by changing the calculation
degrees of freedom and treating the many-body systems as a few-body problem between clusters as done in Ref.~\cite{Arai:2003ek,deDiego:2007rd} introducing approximations in the solution \cite{doweneedthis}.
Resonances in $^4$H have been studied in the past, for example, in Ref.~\cite{deDiego:2007rd} using an effective two-body n-$^3$H potential adjusted to reproduce the $p$-$^3$He experimental phase shifts. 
Here, the resonance energy has been calculated either directly computing the $S$-matrix pole or by performing a $R$-matrix analysis of the phase shifts, showing a sensible difference among the two methods (the $1^-_I$ resonance is located at energy $E_r=1.2$, $\Gamma=3.5$ MeV, and $E_r=3.6$, $\Gamma=5.3$ MeV respectively).
%
% I dont know why this was removed, but it is the main point of the above paragraph
This result further enhances the difference between the position of the S-matrix pole and the resonance position extracted from the cross section, making difficult any kind of comparison between different calculations and experiments.
%
Other calculations have been done in Refs.~\cite{Arai:2003ek,Horiuchi:2013iw,Lazauskas:2019cxj,Li:2021ado} using the resonating group method (RGM) with complex scaling, Spin-dipole strength functions, Faddeev-Yakubowski equations (FYE) in configuration space, and No-core Gamov Shell Model. The results differs from method to method varying in a range $E_R=\{0.9-3.6\}$ MeV and $\Gamma=\{1.0-5.3\}$ MeV.
This variation appears to be a consequence of the different numerical techniques used more than on the interaction employed. In fact, studying the position of the resonance with different realistic interactions but the same theoretical scheme \cite{Lazauskas:2019cxj,Li:2021ado} leads to comparable results.

%
%

In this letter, we study the emergence of a 3+1 $J^{\pi}=1^-$ resonance in $^4$H using LO \eftnopi, demonstrating the possibility of creating n$^3$H $P$-wave states using a purely $S$-wave contact interaction. 
In the region of resonance projectile neutron momentum is $\sim 0.5~m_\pi$, binding momenta of nucleons within $^3$H target are $\sim 0.4~m_\pi$ (for a combined distance from the complete disintegration threshold of $0.8~m_\pi$): thus $^4$H resonant states are at the limit of convergence of \eftnopi, nevertheless still observable in a LO calculation. The absence of any spin-orbit or tensor term in the LO interaction causes degeneracy between $J^{\pi}=0,^-,1^-_I,1^-_{II},2^-$ states which would be splitted including higher orders. 
As $^4$H can still be treated as both a four-body and two-cluster (triton-neutron) system, we choose to employ both the methodologies starting from the same nuclear Hamiltonian.

@@

The theory used in this work is the \eftnopi\cite{vanKolck:1999mw} truncated at LO. The lowest order consists of contact two-body and a contact three-body interactions with corresponding low energy constants (LECs) fitted to reproduce physical observables. To further simplify the theory we truncate it around the SU(4) symmetry \cite{Konig:2016utl}. %the theory is perturbative around su(4) symmetry, so we do not assume but we truncate it
Therefore, $NN$ spin-singlet and spin-triplet interactions are identical and only one contact term appears in the two-body sector. This adds further uncertainty to the theory of order $(a_0m_{\pi})^{-1}\sim30\%$ on the top of the usual \eftnopi LO truncation error. The contact interactions are regularized by an introduction of a Gaussian regulator and a momentum cut-off $\Lambda$. This yields two- and three-body potential 
%
\begin{equation}
    V(\textbf{r})=C_0 \sum_{i,j}^N e^{-\frac{r_{ij}^2\Lambda^2}{4}},
\end{equation}
%
\begin{equation}
    W(\textbf{r})=D_0 \sum_{ijk}^N \left[
    e^{-\frac{(r_{ij}^2+r_{ik}^2)\Lambda^2}{4}}+
    e^{-\frac{(r_{ij}^2+r_{jk}^2)\Lambda^2}{4}}+
    e^{-\frac{(r_{jk}^2+r_{ik}^2)\Lambda^2}{4}}\right],
\end{equation}
%
where the cut-off dependent LECs ($C_0$ and $D_0$) are fitted on two-body observables and triton binding energy ($B_3=8.48$~MeV). 
In order to assess the uncertainty introduced by the adoption of SU(4) symmetry, we use two different parametrizations for the two-body LEC $C_0$: one at unitarity (with $|a_0|>10^5$~fm and referred to as ``unitary'') and one fitting the deuteron binding energy $B_2=2.22$~MeV (referred as ``nuclear''). We assume a degenerate nucleon mass $m=938.858$ MeV.

To extract the effective range parameters and resonance energy we employ and compare results of three different methods. We use two ab initio techniques (four-body calculations): the direct solution of the Faddeev-Yakubovsky equations (FYE) in configuration space \cite{Lazauskas:2004hq, Lazauskas:2019hil} and the stochastic variational methods (SVM)\cite{Suzuki:1998bn}; 
and a potential folding technique that reduces the problem into effective two-body problem, represented by neutron impinging on a solid triton core.

Employing the above interaction we extract the triton-neutron scattering parameters ($a_L$ and $r_L$) for $L=0$ and $L=1$ trough the effective range expansion (ERE) expanded around the triton-neutron threshold:
%
\begin{equation}
    k^{2L+1}\text{cot}(\delta_L)=-\frac{1}{a_L}+\frac{1}{2}r_L k^2 + ... \, ,
    \label{eq.app.ere}
\end{equation}
%
where $\delta_L$ is the triton-neutron phase-shift at angular momentum $L$ and $k$ is the center of mass momentum of the fragments. 
Resonances will, therefore, emerge as poles of the system's $T$-matrix
%
\begin{equation}
T =\frac{4\pi}{m}\frac{1}{\text{cot}(\delta)-i} =
\frac{4\pi}{m}\frac{k^{2L+1}}{\frac{1}{a_L}+...+\frac{1}{2}r_L k^2
-ik^{2L+1}} \label{eq:Tmatrix}
\end{equation}
%
and can be extracted either by fitting phase shifts to the ERE formula or directly through solving the Schr\"odinger equation for complex energies.
%
From an theoretical point of view, extracting the resonance energy with the two methods is not equivalent. 
Solving the Schr\"odinger equation for complex energies is affected by the \eftnopi LO truncation error ($\Gamma_{\eftnopi}\approx k_{res}/m_\pi\sim80\%$), while extracting the resonance position using the ERE of the triton-neutron phaseshift and using RGM formalism (see below) is equivalent to the halo-EFT expanded around t-n treshold.
In the latter case, a new scale, the breaking momentum of $^3$H, is introduced at momentum $K_{d-n}\simeq94$ MeV. 
Since the scattering parameters are extracted using \eftnopi at LO we expect the two methods to agree inside the truncation error $\Gamma_{Halo}\approx k^*_{res}/K_{d-n}\simeq70\%$ using only the scattiring volume and $\Gamma_{Halo}^2$ %should be square, but if someone recheck is better
if the effective range $r_1$ is also included in the ERE.

\vspace{2mm}

%================================================
\subsection{Faddeev-Yakubovsky equation formalism}
%================================================

This formalism, as well as our employed numerical techniques, have been described in detail in Refs.~\cite{Lazauskas:2004hq,Lazauskas:2019hil}.
Here we will just stress that, in order to determine n-$^3$H scattering lengths (volumes), we solve the differential FYE equations by requiring the systems wave function to satisfy boundary condition in the far asymptote for retrieving neutron, defined by:
%
\begin{equation}
\begin{split}
\Psi(X_{^3{\rm H}},\vec{y}_{n-^3{\rm H}})=&\lim_{k\rightarrow0}\psi_t(X_{^3{\rm H}})\big([j_L(k|y_{n-^3{\rm H}}|)/k^{L}]\\
&+a_L(k) [n_L(k|y_{n-^3{\rm H}}|)k^{L+1}]\big)Y_L(\hat{y}_{n-^3{\rm H}});  \label{eq:FYe_a}
\end{split}
\end{equation}
%
here $j_L(z)$ and $n_L(z)$ are spherical Bessel functions of the first and second kind respectively. 
$\psi_t(X_{^3{\rm H}})$ is the wave function of the $^3$H ground state, which is found before undertaking n-$^3$H scattering calculations. 
In the particular case of determining the scattering lengths (volumes) $a_L$, calculations are performed for vanishing momenta $k$.
Note that, in the limit $k\rightarrow0$, the expressions in the brackets are finite functions depending on a distance $|y_{n-^3{\rm H}}|$.

To determine positions of the resonant states we solve FYE by imposing slightly different boundary condition for neutron asymptote:
\begin{equation}
\begin{split}
\Psi(X_{^3{\rm H}},\vec{y}_{n-^3{\rm H}})=&\psi_t(X_{^3{\rm H}}) \big(h_L^+(k_{r}|y_{n-^3{\rm H}}|)\\
&+F_L(k_r)h_L^-(k_{r}|y_{n-^3{\rm H}}|)\big) Y_L(\hat{y}_{n-^3{\rm H}});  \label{eq:FYe_res}
\end{split}
\end{equation}
here $h_L^-(z)$ and $h_L^+(z)$ are Spherical Bessel functions of the third kind. Complex momentum plane $k_r$ is scanned to find positions where the function $F_L(k_r)$ vanishes and thus wave function $\Psi(X_{^3{\rm H}},Y_{n-^3{\rm H}})$ represents purely outgoing n-$^3$H waves.

%================================================
\subsection{Harmonic oscillator trap}
%================================================
The second microscopic approach involves an introduction of the Harmonic oscillator (HO) trapping potential
\begin{equation}
    V_{\rm HO}(\textbf{r}) = \frac{m}{2A} \omega^2 \sum_{i<j} (\textbf{r}_i - \textbf{r}_j)^2
\end{equation}
with oscillator frequency $\omega$ into $A=4$-body $^4$H system.

Assuming that the range of an interaction is much smaller than the typical HO trap range $R << b_{\rm HO} =\sqrt{2/(m \omega)}$ one can match asymptotic $n ^3$H part of a trapped $^4$H wavefunction to the trapped solution of effective $n ^3$H two-body system with no nuclear interaction considered. The $n ^3$H phasehifts $\delta_L$ at relative momentum $k$ are then extracted using generalized Bush formula for an arbitrary orbital momentum $L$ \cite{Suzuki:2009}

>>>>

\author{Johannes Kirscher}
\address{Department of Physics, SRM University - AP, Amaravati 522502, Andhra Pradesh, India}
\address{Theoretical Physics Division, School of Physics and Astronomy,\\
  The University of Manchester, Manchester, M13 9PL, UK}

@@

Elastic neutron-triton scattering is considered at leading order in pionless effective field theory.
In the infinite-cutoff limit, we obtain the scattering length $a_0(\infty) = 3.06(1)$~fm
and the effective range $r_0(\infty) = 1.59(7)$~fm ($S$-wave). 
In the $P$-wave channel, we predict a scattering volume $a_1(\infty) = -9.8(1.6)$~fm$^3$
and a range equivalent $r_1(\infty) = 1.3(4)$~fm$^{-1}$ with the same zero-range theory. 
%
The results are extracted with three numerical techniques: a
Gaussian-stochastic-variational method, an integration of the configuration-space Faddeev-Yakubovsky 
equations, and a resonating-group reduction of the four-body problem to an effective two-body cluster theory. 
Renormalization-group dependence is assessed through variation of a cutoff regulator parametrized in a
range between $1\,\text{fm}^{-1}$ and $10\,\text{fm}^{-1}$.
%
Most remarkably, we find a cutoff-stable/RG-invariant resonance in the $^4$H $J^\pi=1^-$ system.
This $P$-wave resonance is thus a universal consequence of the unitary deuteron system
and an ensuing discretely scale invariant three-body system with a scale set by the triton binding energy.
%
This stabilization of a resonant state in a few-fermion system solely through pure contact interactions
has a significant consequences for the powercounting of the pionless theory.
Specifically, the existence of a resonance implies that the known inability of the theory's leading order
to bind larger nuclei like 16-oxygen can be cured in perturbation theory.

\end{abstract}

@@

Contact effective field theory (EFT) is not only practical in describing
bosonic~\cite{Bazak:2016wxm,Carlson:2017txq,Bazak:2018qnu,Bedaque:1998kg,Bedaque:1998km} 
and few-nucleon~\cite{Bedaque:1999ve,Barnea:2013uqa} systems, it represents a systematic
expansion of the underlying scales and their interactions~\cite{Hammer:2019poc,Barnea:2013uqa}, and
thereby provides a transparent framework to study the emergence of many-body phenomena
from few-body properties.
However, the nuclear incarnation of such a theory, pionless effective field
theory (\eftnopi), apparently fails to predict
stable bound states for $P$-shell
nuclei\footnote{Fermi systems with more particles than accessible internal states}
at the leading order (LO)~\cite{Stetcu:2006ey,Contessi:2017rww,Dawkins:2019vcr,Schafer:2020ivj}. 
%
For instance, the lightest of such nuclei, $^6 {\rm Li}$, is 
found unstable with respect to breakup into $\rm ^4He$ in its ground state ($\alpha$)
and a deuteron ($d$) in the zero-range contact limit~\cite{Schafer:2020ivj}~of LO~\eftnopi.
Whether or not the theory finds the physical bound state at LO veiled as a virtual or resonant state which
reveals its physical character at higher orders is arguably the most imminent problem of the
development of contact EFTs for Fermi systems.
As perturbative higher-order insertions cannot create poles {\it ex novo},
the conceivable absence of such poles
within the convergence radius of LO~\eftnopi~would make the theory in its current version
useless for the description of larger nuclei.
The challenge cast in simple words: 
Does the LO~\eftnopi~support any type of non-zero poles at complex momenta below its breakdown scale?

This issue has been addressed earlier for spatially symmetric states.
Most prominently, a tower of so-called Efimov resonances is predicted in (A$\le$4)-body bosonic systems 
when they approach the unitary limit~\cite{Deltuva:2012ig,deltuva2020energies}. 
Even resonances corresponding to negative $S$-wave effective ranges have been found
compatible with the zero-range theory~\cite{Habashi:2020qgw}.
In the nuclear domain, LO~\eftnopi~finds resonances in the neutral
hypernuclear system $\Lambda nn$~\cite{Schafer:2020rba}
and $S$-wave virtual states in the triton~\cite{Rupak:2019} and hyper triton~\cite{Schafer:2020rba}. 
For the spatially asymmetric wave function of $^3$n, in contrast, neither
a $P$-wave resonance nor a virtual state
were found~\cite{Dietz:2021haj}~cutoff-stable.
In distinction to this latter finding where $^3$n represents a scale-free system,
the subject of our study are channels affected by a finite three-body scale. 
To the best of our knowledge, no simple and convincing argument
can predict the existence of resonances under the aforementioned conditions
solely from this three-body $S$-wave scale. 
The smallest, and experimentally most interesting Fermi system which allows for
a three-particle subsystem to be in a state with totally-symmetric internal wave function,
while the total wave function of at least four particles
must have one of the internal states occupied with two particles, is $^4$H, \ie, the subject of our study. 
This systems acts as a benchmark for the usefulness of~\eftnopi~for a description of even heavier
$P$-shell nuclei, and progress in understanding the nuclear chart beyond the $\alpha$ with
non-perturbative two- and three-body contact interactions as a first approximation must address the
resonant features of $^4$H.

%
Resonances in the $^4$H channel are expressed in n-$^3$H elastic cross sections~\cite{PhysRev.119.1981,SEAGRAVE1972250},
$\pi^-$ absorption experiments~\cite{Sennhauser:1981tr,Gornov:1991fg}, and transfer reactions~\cite{Miljanic:1986zz,Sidorchuk:2003fwa}. 
However, the extracted resonance parameters differ significantly and depend on the used experimental protocol as well as
the underlying hypotheses in the data analysis. 
For instance, from the results of one $\pi^-$+$^9$Be$\,\to\,$d+t+$^4$H
experiment, multiple parameter sets for the lowest state
($E_R=3.0\pm0.2$~MeV,~$\Gamma=4.7\pm1$~MeV)~\cite{Gornov:1991fg}, ($E_R=2.0\pm0.2$~MeV,~$\Gamma=1.2\pm0.2$~MeV)~\cite{Gurov:2005sp},
and ($E_R=1.6\pm0.1$~MeV,~$\gamma^2=0.4\pm0.1$~MeV)~\cite{Gurov2005p}\footnote{$\gamma$ is a so-called reduced width~\cite{Gurov2005p}}
were obtained. These three sets follow from the assumption of one, two, or three resonant states, respectively, being accountable for 
the signal.
An independent, albeit inconsistent R-matrix prediction fed with n-$^3$H cross sections~\cite{Tilley:1992zz}, yields
a superposition of several resonant $P$-wave states with quantum numbers $J^\pi=0^-,1^-_I$, $1^-_{II},2^-$~\cite{Lazauskas:2004uq}~and
groundstate parameters $E_R=3.2$~ MeV and $\Gamma=5.4$~MeV.
%The discrepancies among different experimental approaches further increase when including the transfer reaction experiments.
%

Accurate numerical simulations of resonances in nuclear few-body systems employing system-appropriate interactions are
non-trivial and limited to $A\leq6$~\cite{Li:2019pmg,Lazauskas:2019cxj}.  
For certain channels comprising more nucleons, the usage of fragments which are inert at the considered scattering energies
as effective cluster degrees of freedom allows for practical solutions~\cite{Arai:2003ek,deDiego:2007rd}.
$^4$H resonances, in particular, have been studied in~\cite{deDiego:2007rd}~with an effective two-body n-$^3$H
potential calibrated to p-$^3$He phase shifts. 
The ensuing resonance energy was calculated either as an S-matrix pole or by performing a R-matrix analysis of the phase shifts.
The resultant $1^-_I$ resonance-pole locations differ significantly: 
$E_r=1.2$, $\Gamma=3.5$ MeV, and $E_r=3.6$, $\Gamma=5.3$ MeV, respectively. This discrepancy adds to the one mentioned in the previous
paragraph, and resembles the {\it complexity} of analytically continuing a multi-channel S-matrix with numerous isolated and
cut singularities.
%
% I dont know why this was removed, but it is the main point of the above paragraph
%This result further enhances the difference between the position of the S-matrix pole 
%and the resonance position extracted from the cross section, making difficult any kind 
%of comparison between different calculations and experiments.
%

%
Combining results obtained with the resonating-group method (RGM) and
complex scaling, spin-dipole strength functions, 
Faddeev-Yakubowski equations (FYE) in configuration space, and the
no-core Gamov Shell Model~\cite{Arai:2003ek,Horiuchi:2013iw,Lazauskas:2019cxj,Li:2021ado}
one obtains $E_R=\{0.9-3.6\}$ MeV and $\Gamma=\{1.0-5.3\}$ MeV. The uncertainty is dominated
by numerical approximations while predictions within the same numerical framework are relatively
insensitive to the kind of realistic interaction~\cite{Lazauskas:2019cxj,Li:2021ado}.
%
%

The following report is a first step towards numerical-method and interaction-model-independent 
predictions on the pole structure of $^4$H. While our approach (LO~\eftnopi) has innate uncertainties,
it predicts the existence of a 3+1 $J^{\pi}=1^-$ resonance in $^4$H as an inevitable/universal feature of an almost
unitary system with discrete scale invariance (three-body scale).

% using LO \eftnopi, thus demonstrating the possibility of create n$^3$H $P$-wave states with purely $S$-wave contact interaction. 
%In the region of resonance projectile neutron momentum is $\sim 0.5~m_\pi$, binding momenta of
%nucleons within $^3$H target are $\sim 0.4~m_\pi$ 
%(for a combined distance from the complete disintegration threshold of $0.8~m_\pi$): 
%thus $^4$H resonant states are at the limit of convergence of \eftnopi, nevertheless still 
%observable in a LO calculation. The absence of any spin-orbit or tensor term 
%in the LO interaction causes degeneracy between $J^{\pi}=0,^-,1^-_I,1^-_{II},2^-$ states 
%which would be splitted including higher orders. 
%As $^4$H can still be treated as both a four-body and two-cluster (triton-neutron) 
%system, we choose to employ both the methodologies starting from the same nuclear Hamiltonian.

%% -> to the result section!
We begin with a brief description of the interaction theory (LO~\eftnopi) before introducing the
three numerical techniques which were used to reach our main conclusion. The results leading to that
thesis are presented in the third part prior to a summary.

@@

Nuclear dynamics in the unitary limit, and thereby their universal properties
and correlations with quantum chromodynamics and atomic
physics is described by the~\eftnopi\cite{vanKolck:1999mw}.
From the leading order of this theory, \ie, a contact/zero-range theory,
we deduce our conclusions, and therefore need to introduce its structure. 
It comprises one\footnote{We assume SU(4) symmetry~\cite{Konig:2016utl}~
and thus degenerate singlet and triplet nucleon-nucleon channels.}
two- and a three-body contact interaction, each 
associated with an independent low energy constant (LEC).
%the theory is perturbative around su(4) symmetry, so we do not assume but we truncate it
%Therefore, $NN$ spin-singlet and spin-triplet interactions are identical and only one contact term appears in the two-body sector.
This interaction implies an uncertainty of order $(a_0m_{\pi})^{-1}\sim30\%$.
For the numerical calculations, we find regularization 
with a Gaussian function of width $\propto\Lambda^{-1}$ convenient.
In coordinate representation, the two- and three-body
potentials defined from the bare vertices are:
%
\begin{equation}
    \hat{V}=C_0 \sum_{i<j}^N e^{-\nicefrac{r_{ij}^2\Lambda^2}{4}}
    \quad,
\label{eq:vnn}
\end{equation}
%
\begin{equation}
    \hat{W}=D_0 \sum_{i<j<k}^N \left[
    e^{-\nicefrac{(r_{ij}^2+r_{ik}^2)\Lambda^2}{4}}+
    e^{-\nicefrac{(r_{ij}^2+r_{jk}^2)\Lambda^2}{4}}+
    e^{-\nicefrac{(r_{jk}^2+r_{ik}^2)\Lambda^2}{4}}\right]
    \quad.
\label{eq:vnnn}
\end{equation}
%
The LECs ($C_0$ and $D_0$) are renormalized by fixing the two-body
and triton binding energies.
While the latter is constrained to $B_3=8.48$~MeV, 
we the uncertainty introduced of the SU(4) approximation with two
different conditions on the two-body ground state:
first, we demand an almost zero-energy state with $|a_0|>10^5$~fm to
approximate the unitary limit
(``unitary'' case), and second, we calibrate $C_0$ to the physical deuteron
binding energy $B_2=2.22$~MeV
(``nuclear'' case). All nucleons are assumed to have mass $m=938.858$~MeV.

Effective-range parameters and resonance energies are extracted from solutions of the Schr\"odinger equation, defined with
this potential, with three different methods:
Faddeev-Yakubovsky equations (FYE) in configuration space~\cite{Lazauskas:2004hq, Lazauskas:2019hil}, the stochastic variational method
 (SVM)~\cite{Suzuki:1998bn}, and a potential-folding technique which obtains an effective two-body problem of a neutron scattering off
 an inert triton core.

For the triton-neutron scattering parameters ($a_L$ and $r_L$) in relative $L=0$ and $L=1$ partial waves, we use
the effective range expansion (ERE) about the triton-neutron threshold:
%
\begin{equation}
    k^{2L+1}\cot\delta_L=-\frac{1}{a_L}+\frac{1}{2}r_L k^2 + ... 
    \quad ,
    \label{eq.app.ere}
\end{equation}
%
where $\delta_L$ is the triton-neutron phase shift at angular momentum $L$, and $k$ is the relative momentum of the fragments in the center-of-mass frame. 
Resonances of the system's $T$-matrix are found as complex roots of the denominator of
%
\begin{equation}
T =\frac{4\pi}{m}\cdot\frac{1}{\cot\delta-i} =
\frac{4\pi}{m}\cdot\frac{k^{2L+1}}{-\frac{1}{a_L}+...+\frac{1}{2}r_L k^2
-ik^{2L+1}}
\quad.
\label{eq:Tmatrix}
\end{equation}
%

From the FYE (see~\eqref{eq:FYe_res}~below) we obtain the amplitude directly for complex momenta and
search for its poles near the axes. In contrast, the SVM and the RGM calculate $\cot\delta$ for number of real
momenta approaching $0~$MeV. This series of points is fit to the ERE~\eqref{eq.app.ere} up to second order.
Resonances are then identified with those roots of the polynomial $-ik^{2L+1}$ which lie in the third and forth quadrant
of the complex energy sheet.
%and can be extracted either by fitting phase shifts
% to the ERE formula or directly through solving the Schr\"odinger equation for complex energies.
%
The uncertainties in the predicted pole locations and ERE parameters thus differ from method to method. 
%From an theoretical point of view, extracting the resonance energy with the two methods is not equivalent.
The theoretical uncertainty in the SVM and FYE calculations is expected to decrease with the convergence rate
of the pionless theory, namely, $\approx k_{\rm typ}/m_\pi\sim80\%$ with a typical momentum set by the
resonance location $k_{\rm res}$ (see results).
%Solving the Schr\"odinger equation for complex energies is affected by the
% \eftnopi LO truncation error ($\Gamma_{\eftnopi}\approx k_{res}/m_\pi\sim80\%$),
%  while extracting the resonance position using the ERE of the triton-neutron
%   phaseshift and using RGM formalism (see below) is equivalent to the halo-EFT expanded around t-n treshold.
The breakdown scale, and thus the convergence rate of the RGM is set by the lowest excitation threshold of $^3$H.
%In the latter case, a new scale, the breaking momentum of $^3$H, is introduced at momentum 
We associate a momentum $k_{d-n}\equiv\sqrt{\nicefrac{4}{3}\, m\,(B_3-B_2)}\simeq94$ MeV with this excitation.
With this smaller area of applicability of the cluster/RGM formulation, a pole predicted at some $k_{\rm res}\approx k_{d-n}$
with the SVM or FYE, whose breakdown is around $m_\pi$, might not be seen with the RGM. In fact, this difference in convergence radii
along with the relatively large uncertainty of $\sim80\%$ explains the results which we will discuss
after the following details on the numerical methodology.

%Since the scattering parameters are extracted using \eftnopi at LO
% we expect the two methods to agree inside the truncation error
% $\Gamma_{Halo}\approx k^*_{res}/K_{d-n}\simeq70\%$ using only
% the scattiring volume and $\Gamma_{Halo}^2$ %should be square, but if someone recheck is better
%if the effective range $r_1$ is also included in the ERE.

%================================================
\subsection{Faddeev-Yakubovsky-equation formalism}
%================================================

The Faddeev-Yakubovsky formalism decomposes the few-body Schr\"odinger/Lippman-Schwinger equation into a
set of equations in order to organize the overlapping spectra and divergences of the few-body scattering 
problem. Details about this complex and on our particular numerical realization can be found in
Refs.~\cite{Lazauskas:2004hq,Lazauskas:2019hil}.
The appropriate boundary conditions on the formalism which yield the n-$^3$H scattering lengths (volumes)
relevant for this work,
%
\begin{equation}
\Psi(t,\ve{y})=\lim_{k\rightarrow0}\,\left(\psi_t\,\left[j_L(ky)/k^{L}+a_L(k)\, n_L(ky)k^{L+1}\right]\,Y_L(\hat{\ve{y}})\right)
\quad,  \label{eq:FYe_a}
\end{equation}
%
fix the solutions of the differential FYE equations in the far asymptote where the radial distance $\left|\ve{y}\right|$
between the neutron and $^3$H is larger than the support of $\hat{V}$ and $\hat{W}$.
Here, $j_L(z)$ and $n_L(z)$ are spherical Bessel functions of the first and second kind,
respectively, and 
$\psi_t$ is the $^3$H ground-state wave function whose determination precedes n-$^3$H scattering calculations. 
wERE parameters are extracted in the limit $k\to0~$MeV.
In this limit, both terms in the bracket remain finite functions of a distance $|y_{n-^3{\rm H}}|$.

Resonant states are obtained by imposing the canonical boundary conditions on the FYE, namely a purely outgoing wave
in the region where the neutron projectile is no longer interacting with any target $^3$H nucleons:
\begin{equation}
\Psi(t,\ve{y})=\psi_t\,\left[h_L^+(k_{\rm res}y)+F_L(k_{\rm res})h_L^-(k_{\rm res}y)\right]\,Y_L(\hat{\ve{y}})
\quad.  \label{eq:FYe_res}
\end{equation}
Now, $h_L^-(z)$ and $h_L^+(z)$ are Spherical Bessel functions of the third kind, and 
the solution corresponds to a resonant state for those complex momenta $k_{\rm res}$ at which the
function $F_L(k_{\rm res})$ vanishes, \ie, where the wave function $\Psi(t,\ve{y})$ represents
a purely outgoing n-$^3$H waves.
